On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2017-09-22 manuscript.Rmd
Text as of draft zero - sent to coauthors

```{r include=FALSE}
includefigs <- F
library(png)
library(grid)
library(knitr)

@@

knitr::kable(tab[[1]], col.names="")

@@

Stomata exert a strong control on the transpiration rate of leaves. During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (gmin). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity). It should be clear from Eq. 1 that the Ball-Berry model is actually a model of water-use efficiency, since both g~s~ and A~n~ appear in the equation. Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored. For example, a recent review of stomatal conductance models did not mention the minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. Discuss @barnard2013 briefly (in a positive way :)

\
(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models. We aim to improve the connection somewhat, and point at obvious gaps in understanding. Throughout our review, we discuss both the practical implementation of a 'minimum' conductance g~0~ as implemented in e.g. Ball-Berry models of stomatal conductance (Eq. 1) and similar, and a 'true' minimum conductance as estimated on detached leaves (g~min~).

@@

Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001] (g~cut,isol~), or via gas exchange techniques, or mass loss of detached leaves vby sealing the side with stomata (g~cut,seal~). The method we here refer to as minimum conductance is that of mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. The g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

\ 
We quantitatively compared each of these different 'minimum' codnductance values by synthesizing existing data. To estimate g~0~ from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.  Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo A for a list of papers and additional figures describing this new compilation). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

@@

@@

The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be concluded from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values.

@@

\
In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated. We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied. Recently, @barnard2013 also mention the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details.

\
In Fig. 2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study. The correlation between slopes and intercepts is not unique to Eq. 1, but a general property of linear models [@stapleton1995]. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts).

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and m (applying the model of ) from @miner2016. In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because environmental conditions varied little. We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation the ball-berry coefficient (i.e. right-hand side of Eq. 1) is lower (Fig. 3c).

@@

\
The effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers. The g~0~ parameter not only sets a minimum value of g~s~ in the model, it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata [@kerstiens1996]. Finally, measuring the low water loss rates of detached leaves mimicks drought conditions, yielding an estimate of whole-leaf minimum conductance (g~min~) [@pisekberger1938; @hygen1951]. Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in concert [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~.

@@

\
The g~0~ parameter also represents g~s~ when T~leaf~ is very high, because A~n~ goes to zero at high T~leaf~. As a result, simulations of vegetation water use during heat waves depends highly on the value of g~0~ used [@kala2016] . It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).

\
Using the coupled leaf gas exchange model, we show how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1.

@@

Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000].

Various new models that aim to simulate desiccation have a g~min~ term in it, but it is not yet well constrained by data.

@@

Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance, even compared to measurements based on MLD with stomatal side sealed (g~mld,seal~). The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated as they are relevant to the current discussion.

@@

\
Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata. Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.

@@

A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~). However as our literature comparison showed, g~dark~ is substantially higher compared to g~s~ reached at low photosynthetic rates, and g~min~. Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same species [@howard2007; @caird2007]. The idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

\
First, ample observations suggect that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.

@@

Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved. From this minimum transpiration rate, a minimum conductance (g~min~) is usually estimated, taking into account the vapour pressure deficit [@sinclair1986]. This method has been termed the 'mass loss of detached leaves' (MLD) [@kerstiens1996]. Many previous reports use the term 'cuticular conductance' for g~min~, but this is inaccurate as some fraction of the water loss will occur through broken, leaky, or incompletely closed stomata. Our literature review was somewhat inconclusive in this regard, as conductance g~min~ measured on leaves where the stomatal side was sealed was similar (across all included studies) to g~min~ of whole leaves (Fig. 3). More conclusive is evidence from individual studies where the stomatal component to g~min~ could be quantified. A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten Sorghum genotypes, @muchow1989 reported a correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky. However @santrucek2004 found that the cuticle (solid mass only) of the stomatal side in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles.

@@

In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants. We also further tease apart the literature compilation of g~min~ (shown just as a single average value in Fig. 4).

Using the literature compilation g~min~ (see SectionX), we asked whether g~min~ shows clear variation due to, species growth form, taxonomic order, and average climate across the species' distribution.

We found not so much! Grasses have a higher gmin. So do crops, and within crops there is a ton of genotypic variation.

Very few studies have found meaningful correlations. Exception is @brodribb2014, showing a correlation between rainfall at species origin (of the driest quarter), and g~min~.

Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~ (see SuppInfo 2).

@@

knitr::kable(dat, 
             col.names= c("Study","Species","Experiment","Reduction in g~min~"))
```

>>>>

```{r include=FALSE}
includefigs <- T
library(png)
library(grid)
library(knitr)
library(pander)

@@

knitr::kable(tab, col.names="")

@@

Stomata exert a strong control on the transpiration rate of leaves. During severe drought, or conditions where photosynthesis does not take place, it seems advantageous for plants to completely halt water loss. However, water loss continues through the cuticle and incompletely closed stomata, together determined by the leaf minimum conductance (g~min~). In this review, we illustrate the role of g~min~ in models of leaf gas exchange, critically evaluate sources of variation in g~min~, and quantitatively compare various methods for its estimation. A literature compilation of g~min~ as measured by weight loss of detached leaves is presented, which shows much variation in this trait that is not clearly related to species groups, climate of origin, or leaf type. Much evidence points to the idea that g~min~ is highly responsive to the growing conditions of the plant, including soil water availability, temperature, and air humidity - as we further demonstrate with two case studies. The synthesis presented here provides guidelines for use of g~min~ in ecosystem models, and points to clear research gaps for this drought tolerance trait.

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, and f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987), and g~0~ is the value of g~s~ when A~n~ is zero. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987) (they used relative humidity). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been largely ignored. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is estimated from a regression of Eq. 1 with leaf gas exchange data, but this may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

\
(*expand*) We thus have several different definitions of what constitutes a 'minimum' conductance of leaves. In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values to the minimum conductance. We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different.

\
(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple  assumptions made when applying ecosystem models. We aim to improve the connection somewhat, and point at obvious gaps in understanding.

@@

\
Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001] (g~cut,isol~), or via gas exchange techniques, or mass loss of detached leaves vby sealing the side with stomata (g~cut,seal~). The method we here refer to as minimum conductance is that of mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. The g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

\ 
We quantitatively compared each of these different 'minimum' conductance values by synthesizing existing data. To estimate g~0~ from regression, we used the @lin2015 database to fit the model of @medlyn2011 using non-linear regression, and data presented by @miner2017 on the linear regression of the @ballberry1987 model. Measurements of conductance of isolated, non-stomatal cuticles (g~cut,isol~), and of conductance of detached leaves with the stomatal side sealed (g~cut,seal~) were taken from the compilation by @kerstiens1996.  Data on minimum conductance (g~min~) from MLD was compiled from 44+ original papers (see SuppInfo A for a list of papers and additional figures describing this new compilation). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources. All conductance values were converted to per unit projected surface area. Units were converted to molar fluxes assuming 41 mol m^-3^ (cf. @kerstiens1996). Although it is quite clear that night-time conductance (g~dark~) is much larger than a true 'minimum' conductance, it is frequently used as an independent measurement of g~0~ [@lombardozzi2017], and thus a comparison of magnitudes is useful. A compilation of night-time conductance (g~dark~) values was taken from @lombardozzi2017, who compiled cuvette (and sometimes sap flow-based) measurements of ca. 150 species during night-time. Actual minimum daytime values were estimated from the Lin et al. database, by taking those data where A~n~ < 2 $mu$ mol m^-2^ s^-1^, and calculating the minimum g~s~ for each of the XXX studies included in that dataset. This estimate was added as a reference as it represents an actual minimum values measured, rather than a fitted/extrapolated g~0~ term.

@@

```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure1.png")
```

**Figure 1**. Comparison of various estimates of the minimum conductance. Error bars are 95/% confidence intervals. See Table 1 for the various definitions. Different letters denote significant differences (alpha = 0.05). Note the logarithmic scale.

**In case you are interested, this is what the same graph looks like with no log scale**

```{r fig.width=4,echo=FALSE, eval=includefigs}
plotfig("output/figure1_unlog.png")
```

**Figure 1 Version 2**. As Figure 1, but natural scale.

@@

\
The standard approach in estimating g~0~ for use in Eq. 1 is to use regression with g~s~ as the response variable, and the right-hand side of Eq. 1 as the predictor. The g~0~ parameter is thus estimated as the intercept. Here we briefly discuss some statistical aspects of this estimation procedure, and draw the general conclusion that g~0~ is poorly estimated by this method. The difficulty of the estimation can already be anticipated from the fact that a) many studies set g~0~ to some assumed value rather than fitting it (e.g. @leuning1995 uses 0.01 mol m^-2^ s^-1^ for all species); presumably because unreliable estimates were obtained, b) negative g~0~ estimates from regression are commonly reported [@leuning1995; @heroult2013; @miner2017], though it is clearly non-sensical to suggest negative conductance values. @barnard2013 also mentioned the difficulty of fitting Eq. 1 to estimate g~0~, but did not present specific details.

@@

\
In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated. We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.

\
In Fig. 2, we demonstrate how g~0~ and g~1~ are statistically correlated, when fitted for a single study. The correlation between slopes and intercepts is not unique to Eq. 1, but a general property of linear models [@stapleton1995]. The consequence is that it is more difficult to interpret g~1~ and g~0~ independently, and that confidence intervals on each of the parameters should also not be interpreted independently (see @becker2007 for more discussion on the statistical aspects of comparing slopes and intercepts). (*expand*)

**Figure 2**. Example linear regression of stomatal conductance (g~s~) against a stomatal index, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of Martin-StPaul (cite). This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

\
Next we study the values obtained when fitting Eq. 1 to many datasets, basing our work on the large database collected by @lin2015, and the compilation of fitted g~0~ and m (applying the model of ) from @miner2017. In both cases, estimates of g~0~ are inflated when the model fits poorly (Fig. 3 a,b). This can be understood by considering that a poor fit often results in the flattening out of the regression line, thus giving a large value for the intercept. A poor fit is often obtained when there is little variation in the right-hand side of Eq. 1, usually because environmental conditions varied little. We confirm this by showing that the standard error (SE) of g~0~ increases when the coefficient of variation the ball-berry coefficient (i.e. right-hand side of Eq. 1) is lower (Fig. 3c).

**Figure 3**. (*Might go in SuppInfo?*) Correlation between estimated g~0~ and the R^2^ of the regression. In the left panel, we fitted the linearized form of the Medlyn et al. (2011) model to each of the datasets in the Lin et al. (2015) database. Vertical lines are 95% confidence intervals. In the right panel, a very similar association between g~0~ and R^2^ is presented by the review of @miner2017, using the Ball-Berry model. In panel (c) coefficient of variation and R2.

@@

\
In the Ball-Berry type framework of stomatal conductance models, the effects of photosynthetic photon flux density (PPFD) and leaf temperature on g~s~ are both assumed to enter via the dependence of A~n~ on these drivers. Here we show that the g~0~ parameter not only sets a minimum value of g~s~ in the model (when A~n~ is zero), it also modifies the behaviour of Eq. 1 after it has been coupled to the Farquhar-von Caemmerer-Berry model of photosynthesis (the so-called 'coupled leaf gas exchange model'). @leuning1990 defined g~0~ as g~s~ when A approaches zero as PPFD approaches the light compensation point, and pointed out that g~0~ is necessary to simulate the increase in A/g~s~ (and C~i~) at low light [see also @leuning1995]. Similarly, @collatz1992 showed that including g~0~ affects the response of g~s~ to relative humidity in the coupled leaf gas exchange model. Despite these early reports, the exact role of g~0~ in models is often overlooked.

\
Within the model framework, g~0~ is reached when photosynthesis (A~n~) goes to zero. But A~n~ can approach zero for many different reasons, including low light, high temperature, low humidity, and drought, begging the question whether the same g~0~ is reached in each of these cases. In the application of the Ball-Berry models, it is frequently assumed that g~0~ must equal g~s~ at night, simply because no photosynthesis occurs at night [@lombardozzi2017; @barnard2013; @uddling2005]. It is also frequently assumed that g~0~ is equal to the absolute minimum conductance achievable for a leaf, the 'cuticular conductance' (g~cut~) [@baldocchi1997; @manzoni2011; @egea2011], which technically is the conductance of the cuticle alone - ignoring leaky and incompletely closed stomata (see Section XXX).

\
Photosynthesis also ceases at very high temperature (ref), but stomata do not appear to close in proportion to the decrease in photosynthesis, if at all [@drake, @urban2017], thus we cannot use high temperature to estimate a reasonable g~0~. Simulations of water use during heat waves depends highly on the value of g~0~ used [@kala2016] . It should be pointed out that the response of g~s~ to T~leaf~ is likely poorly modelled with a Ball-Berry model, since a number of studies have shown that g~s~ and A~n~ decouple at high T~leaf~, resulting in high rates of g~s~ when A~n~ is near zero [@urban2017] (Drake et al., unpublished).

@@

\
Using the coupled leaf gas exchange model, we show in Fig. 4 how A~n~/g~s~ and C~i~ depend on PPFD with three values of g~0~ (0, 0.01 and 0.03 mol m^-2^ s^-1^). We also demonstrate the effect of g~0~ on modelling the T~leaf~ response. In this simulation, T~leaf~ and D are assumed to co-vary with an empirical relationship as used by @duursma2014. Over the entire range of T~leaf~, inclusion of a non-zero g~0~ obviously increase leaf transpiration E~L~, but at a slightly higher rate than just due to g~0~. This effect arises because the additional conductance allows slightly higher rates of photosynthesis, which in turn increase g~s~ via Eq. 1. (*expand*, cite some papers showing these are the responses we expect).

```{r fig.height=6,echo=FALSE, eval=includefigs}
plotfig("output/figure4.png")
```

**Figure 4**. Simulations  of intrinsic water-use efficiency (A~n~/g~s~), intercellular CO~2~ concentration (C~i~) and leaf transpiration rate (E~L~) with three values of g~0~. For the top two panels, all environmental drivers were held constant except the photosynthetic photon flux density (PPFD), demonstrating that the inclusion of g~0~ is necessary to simulate the decrease in A~n~/g~s~ at low PPFD. For the bottom two panels, VPD and T~air~ were covaried based on an empirical relationship [@duursma2014], reflecting typical covariation in field conditions. For all simulations it is assumed that T~leaf~ is equal to T~air~.

@@

\
Another reason to include a non-zero g~0~ term in models of stomatal conductance arises during severe drought. When plants are sufficiently water-stressed that stomata are completely closed, water loss still continues through the cuticle and leaky and incompletely closed stomata. Thus models that aim to predict when plants desiccate and die must include a minimum conductance term [@blackman2016; @sinclair2000]. Various new models that aim to simulate desiccation have a g~min~ term in it (even when not explicitly defined in the main equations) [@xu2016] (sperry, others), but it is not yet well constrained by data.

@@

\
Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance (Fig. 1), in line with previous comparisons [@kerstiens1996]. The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. Indeed, some more recent studies have shown a very good correspondence between g~cut~ and g~min~ (riederer, schuster). A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.

@@

\
(*expand*) Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata (see Section XXX). Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions.

@@

\
A number of studies have assumed that the g~0~ in the Ball-Berry model equals the night-time conductance (g~dark~) [@barnard2013; @lombardozzi2017], simply because it is a condition where no photosynthesis takes place. However as our literature comparison showed (Fig. 1), g~dark~ is substantially higher compared to any other measure of minimum conductance. Previous work on g~dark~ have confirmed that night-time conductance is generally much higher than g~min~, when measured on the same plants [@howard2007; @caird2007; @walden-coleman2013; @phillips2010]. The simple idea that g~dark~ is equal to the minimum reached during periods of near-zero PPFD or high temperature during the day can be dismissed based on several observations.

\
First, ample observations suggest that g~dark~ is not a fixed rate but varies tremendously during the night. @caird2007 (and references therein) described how for many species, g~dark~ is not stable throughout the night period. Instead, endogenous, gradual increases in stomatal opening during predawn hours have been reported in many species under natural field conditions as well as in controlled environments [@rawson1988]. @rescodedios2016 showed that, in *Eucalyptus camaldulensis*, g~dark~ in the period just after sunset was much lower than pre-dawn g~dark~. Strong evidence for endogenous regulation of g~dark~ was reported by @rescodedios2013, who showed that g~dark~ fluctuated throughout the night despite environmental conditions being held constant in whole-tree chambers. In addition, g~dark~ can show a clear response to vapour pressure deficit [@barbour2007] during night-time conditions. Clearly g~dark~ is an actively controlled process that cannot be adequately summarized by a single constant g~0~.

@@

Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after a while a constant low rate is achieved. From this minimum transpiration rate, a minimum conductance (g~min~) is usually estimated, taking into account the vapour pressure deficit [@sinclair1986]. This method has been termed the 'mass loss of detached leaves' (MLD) [@kerstiens1996]. Many previous reports use the term 'cuticular conductance' for g~min~, but this is inaccurate as some fraction of the water loss will occur through broken, leaky, or incompletely closed stomata. A few studies have directly quantified the stomatal component of g~min~. A detailed study of *Hedera helix*  concluded that 35\% of water loss occurred across the stomatal pores (though closed), and 65\% across the non-stomatal part of the cuticle, despite the fact that the stomata presumably covered only a small fraction of the leaf [@santrucek2004]. For ten Sorghum genotypes, @muchow1989 reported a strong positive correlation between stomatal density and g~min~, suggesting leaky stomata contribute substantially to g~min~. For seven conifer species, @brodribb2014 reported that 50 - 94\% of water loss of detached leaves originated from the stomatal side, concluding that stomata must be very leaky. However @santrucek2004 found that the cuticle of the stomatal side (but excluding stomata itself) in *Hedera helix* was many times more more permeable than the non-stomatal side, thus providing another potential explanation for this difference. Either way, water loss rates of intact leaves cannot be simply explained from the permeability of isolated non-stomatal cuticles, and the contribution of leaky stomata, and potentially more permeable cuticle on the abaxial side, likely varies between species.

@@

In the following sections, we review the quite substantial literature on g~min~ measured with the MLD technique, stretching back to the 1930s [@pisekberger1938], focussing on environmental and ecological determinants. We also further tease apart the literature compilation of g~min~ (which was shown just as a single average value in Fig. 4).

We found it difficult to explain the variation in g~min~ between species. Grasses have a higher g~min~, and so do crops. Otherwise few differences were significant, with Pinales tending towards lower values of g~min~, and Myrtales (including many *Eucalyptus* species), tending towards higher values.

Comparing all available data, we did not find significant relationships between rainfall, aridity, or temperature and g~min~ (see SuppInfo 2). There was a weak correlation between g~min~ and P~50~, but this was driven entirely by the difference in P~50~ between gymnosperms and angiosperms (not shown).

Very few studies have found meaningful correlations between g~min~ and environmental factors or ecological categories. One notable exception is @brodribb2014, who showed a correlation between rainfall at species origin (of the driest quarter), and g~min~.

**Figure 5**. Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves. Left panel, probability density function of all values (after averaging by species, n=200+), with a normal distribution curve (mean=, sd=, on the log scale). Middle panel, g~min~ averaged by phylogenetic order (including only the top ten orders in the database). Bars are 95/% confidence intervals. Numbers on top refer to the number of species, different letters denote significant differences (at alpha = 0.05, adjusted for multiple comparisons), and grey symbols are individual species. Right panel, data for crops only where each point is a genotype. Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species. References for panels A and B can be found in the SuppInfo 1, references for panel C are XXX. (*have to add some labels*).

@@

knitr::kable(dat, 
             caption="**Table 2.** Summary of studies demonstrating a decrease in g~min~ after an acclimation period to low soil water availability",
             col.names= c("Study","Species","Experiment","Reduction in g~min~"))
```
